---
title: Time Series Analysis on GE Stock Data to Predict the Close Price of a Stock in
  FUTURE
output:
  html_document: default
  html_notebook: null
  pdf_document: default
  theme: united
  toc: yes
  toc_float: yes
---
### Problem Description
 
 Build time series models in using GE Stock data and the goal is to predict the closing price of a
certain stock for future months ahead using time series algorithms and evaluating each of the
algorithms.


###Clear the Global Environment
```{r}
rm(list=ls(all=TRUE))
```
### Library Call
```{r}
library(forecast)
library(lubridate)
library(DataCombine)
library(imputeTS)
library(dplyr)
library(TTR)
library(graphics)
library(data.table)
library(plyr)
library(zoo)
```
### Read Data from csv
* read the csv file
```{r}
data=read.csv("C:/insofe\\lab_assignments\\Time series\\bse_ge_power_stock.csv")
```
### Observe the data
* What are all the columns given in the data
```{r}
names(data)
##Summary
summary(data$Close)
## Dimensions
dim(data)
## Unique
summary(data$Date)
```
## Basic data View
```{r}
head(data,30)
```
## Summary of the entire data set
```{r}
summary(data)
str(data)
```

### Observation & Analysis 
* We need to have one data point per month price so therefore we need to aggregate them by day of * month
* R has read the data column as factor we need convert it to date format
```{r}
data2=data.frame(data[,c("Date","Close")])
head(data2)
data2$Date=as.Date(data2$Date,format="%m/%d/%Y")
```
### Missing values in time series
* Some times there will be missing entries in dates which will create a missing day in the data or if it is quarter,month or annual .
* Observe the data to find if any
```{r}
head(data2)
str(data2)
```
### Detection of the missing values
* Create a date field which consists of continuous sequence of dates 
* Check against this with the current price data and find out the missing dates.
* Join this variable to the current data to see the missing dates

```{r}
minDate=min(as.Date(data2$Date,format="%m/%d/%Y"))
maxDate =max(as.Date(data2$Date,format="%m/%d/%Y"))
```
### Creating a data frame for sequence of Days
```{r }
seq <- data.frame("dateRange"=seq(minDate,maxDate,by="days"))
```
#### Merge the data frames of seq of dates with the dates of sale 
```{r }
data3=merge(seq,data2,by.x="dateRange",by.y="Date",all.x=T)
head(data3)
```

### Impuation of Missing Values
* Replace the missing values with proceeding value or succeeding or averaging
  both.
* To do that we can use na.locf of "zoo" package which will give succeding values
* we can use rev( ) to reverse the sequence and take the average of both
```{r}
data3$Close<-na.locf(data3$Close)
head(data3)
```

### Observation on MEAN_PRICE
* We need to do monthly aggregation instead of daily since we have 10 years of data
* Adding Month column in addition can help in this 
```{r}
data3$Month <-as.numeric(format(data3$dateRange,format="%Y.%m"))
data4<-ddply(data3, .(Month),summarize,MEAN_Close=mean(Close))
head(data4)
```

### Splitting of the Data
* Random split is not possible because here we will be in need of sequence where by we miss the data points
* splitting is done by sequential splitting
```{r}
Train=data4[which(data4$Month<=2017.04),]
Test=data4[which(data4$Month>2017.04),]
```

### converting into time series 
* our target variable is close price and each month has a close price aggregated
```{r}
Price <- ts(Train$MEAN_Close,frequency =12)
```

### Vizualize the time series Data
```{r}
plot(Price,
     type="l",
     lwd=3,
     col="red",
     xlab="yearly",
     ylab="Price",
     main="Time series plot for Close price")
```
### Decomposed Time Series
* Decompose will provide us with the info on seasonality,trend and randomness
```{r}
Pricedecomposed=decompose(Price)
plot(Pricedecomposed,col="Red")
```
### ACF,PACF 
* Autocorrelation is the linear dependence of a variable with itself at two points in time
* For stationary processes, autocorrelation between any two observations only depends on the time lag h between them
*  Partial autocorrelation is the autocorrelation between yt and yt–h after removing any linear dependence on y1,y2, ..., yt–h+1
```{r}
par(mfrow=c(2,2))
acf(Price,lag=30)
pacf(Price,lag=30)
```


```{r }
Price1 <- ts(Train$MEAN_Close, frequency =1)
plot(Price1,xlab = "Monthly",ylab="Price")
acf(Price1,lag=30)
pacf(Price1,lag=30)
```

Looking at the Y scale in ACF we observe that trend is more dominant than seasonality
* Data is not stationay and we need to stationarize the data

### Stationarize by differencing

```{r}
par(mfrow=c(2,3))
plot(diff(Price1,lag = 1),type="l"); acf(diff(Price1,lag = 1),lag=30) ;pacf(diff(Price1,lag = 1),lag=30)
plot(diff(Price1,lag=2),type="l");  acf(diff(Price1,lag = 2),lag=30); pacf(diff(Price1,lag = 2),lag=30)
```
* one lag has stationarize the data we can use ndiffs of forecast package to check no of differences required to      stationarize the data
```{r}
ndiffs(Price)
```
### Weighted Moving Averages
```{r}
fitwma<- WMA(Price,n=2,1:2)
wmaMape <- mean(abs((Price[2:length(Price)]-fitwma[2:length(Price)])/Price[2:length(Price)]))
wmaMape
pred<-forecast(fitwma,h=4)
plot(pred)
lines(fitwma,col="yellow")
```

### Exponential Moving Averages
```{r}
fitEma <- EMA(Price, n = 2)
emaMape <- mean(abs((Price[2:length(Price)]-fitEma[2:length(Price)])/Price[2:length(Price)]))
emaMape
pred<-forecast(fitEma,h=4)
plot(pred)
lines(fitEma,col="purple")
```


## Build a HoltWinters model  with trend 
```{r}
holtpriceforecast <- HoltWinters(Train$MEAN_Close,gamma=FALSE)
plot(holtpriceforecast)
head(holtpriceforecast$fitted)
```
## HoltWinters model  with trend  and Seasonality

```{r}
priceholtforecast <-
HoltWinters(Price, beta=TRUE, gamma=TRUE, seasonal="additive")
plot(priceholtforecast)
head(priceholtforecast$fitted)
```
*Since we are building the models on monthly data, we will get 12 seasonal components.

### Prediction on the Train
```{r}
holtforecastTrain <- data.frame(priceholtforecast$fitted)
holtforecastTrainpredictions <- holtforecastTrain$xhat
head(holtforecastTrainpredictions)
```

### Prediction on test data
```{r}
priceforecast<-forecast(priceholtforecast,h = 2)
plot(priceforecast,ylim = c(-1000,1000))

```
## Considering the difference from the graph as d=1 to stationarize
```{r}
model1 <- arima(Price,c(0,1,0))
model1
acf(diff(Price,lag = 1))
pacf(diff(Price,lag = 1))
plot(diff(Price))
```
* plot has still non stationary behaviour another difference can stationarize it 
```{r}
model2 <- arima(Price,c(0,2,0))
model2
plot(diff(Price,differences = 2))
acf(diff(Price,differences = 2))
pacf(diff(Price,differences = 2))
```
* Observing the acf and pacf there is significant lag in acf and also in pacf that has to be taken care 
```{r}
model3 <- arima(Price,c(1,1,1))
model3


```
## Plots of the models

```{r}
par(mfrow=c(2,2))
plot(model1$residuals,ylim=c(-500,500))
plot(model2$residuals,ylim=c(-500,500))
plot(model3$residuals,ylim=c(-500,500))
```
###  Auto Arima
```{r}
MODEL_ARIMA <- auto.arima(Price, ic='aic')
summary(MODEL_ARIMA)
```

### Forecast on the models 
```{r}
pricearimaforecasts1 <- forecast(model1, h=2)
plot(pricearimaforecasts1)
pricearimaforecasts2 <- forecast(model2, h=2)
plot(pricearimaforecasts2)
pricearimaforecast3 <- forecast(model3, h=2)
plot(pricearimaforecast3)
pricearimaforecasts_autArima<- forecast(MODEL_ARIMA,h=2)
plot(pricearimaforecasts_autArima,flwd = 2)
summary(MODEL_ARIMA)

```
* Model 1 was constructed with no trend and no seasonality and therefore the prediction will be same as present.

*Model3 has both trend and seasonality.
*Comparing the AIC Auto Arima is giving good results for the future stocks prediction